---
title: "Partial Dependence Plots and their effect on the rate of early recurrence for Elastic Net (ELR), Random Forest (RF), XGBoost and Ensemble models"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
{library(readxl)
library(dplyr)
library(caret)
  library(ggpubr)
  library(plotly)
  library(DiagrammeR)
  library(qs)
  library(plotrix)
  library(tableone)
library(pROC)
  library(doParallel)
library(givitiR)
  library(CORElearn)
library(ResourceSelection)
library(AF)
library(rms)
library(plyr)
library(haven)
library(survival)
library(paf)
library(gridExtra)
library(readxl)
library(haven)
library(mice)
library(VIM)
library(boot)
library(tidyverse)
library(rfUtilities)
library(gbm)
library(DescTools)
library(glmnet)
library(rpart)
library(DescTools)
library(paf)
library(caretEnsemble)
  library(DALEX)
}
intextVal<-qread('intextVal.q')
validationlist<-qread('validationlist4.1.q')
{
  ELRFig<-ggarrange(validationlist$Calibration[[1]]+ggtitle('Uncalibrated'),
            validationlist$Calibration[[5]]+ggtitle('Calibrated'),ncol=2)
  ELRFig<-annotate_figure(ELRFig,
                  top = text_grob("Elastic Net Regression", color = "black", face = "bold", size = 14))
     RFFig<-ggarrange(validationlist$Calibration[[2]]+ggtitle('Uncalibrated'),
                       validationlist$Calibration[[6]]+ggtitle('Calibrated'),ncol=2)
     RFFig<-annotate_figure(RFFig,
                             top = text_grob("Random Forest", color = "black", face = "bold", size = 14))
     XGBFig<-ggarrange(validationlist$Calibration[[3]]+ggtitle('Uncalibrated'),
                       validationlist$Calibration[[7]]+ggtitle('Calibrated'),ncol=2)
     XGBFig<-annotate_figure(XGBFig,
                             top = text_grob("XGBoost", color = "black", face = "bold", size = 14))
     XGBFig    
     
     EnsFig<-ggarrange(validationlist$Calibration[[4]]+ggtitle('Uncalibrated'),
                       validationlist$Calibration[[9]]+ggtitle('Calibrated'),ncol=2)
     EnsFig<-annotate_figure(EnsFig,
                             top = text_grob("Ensemble", color = "black", face = "bold", size = 14))
     }
dataC<-readRDS('dataC4.1.rds')
{
ELGrid = expand.grid(alpha = seq(0,1,by=0.1),lambda = seq(0,0.1,by = 0.005))
RFGrid = expand.grid(
  .mtry=2:10,
  .splitrule=c("gini","extratrees"),
  .min.node.size=c(1,3,5)
)
XGBGrid <- expand.grid(
  nrounds = seq(from = 0, to = 1000, by = 200),
  eta = c(0.05, 0.1, 0.3),
  max_depth = c(2, 4, 6),
  gamma = c(0.5,0.7,1.0),
  colsample_bytree = c(0.4,0.8),
  min_child_weight = c(1,2,3),
  subsample = c(0.5,0.75))

  cctrl1 <- trainControl(method="repeatedcv", number=10,returnResamp = "all", repeats=5, 
                         classProbs = TRUE, summaryFunction = multiClassSummary, savePredictions = TRUE, verboseIter = TRUE,allowParallel = TRUE)
  cctrl3 <- trainControl(method="boot", number=1000,returnResamp = "all", 
                         classProbs = TRUE, summaryFunction = multiClassSummary, savePredictions = TRUE, verboseIter = TRUE)
  cctrl3a <- trainControl(method="boot", number=1,returnResamp = "all", 
                         classProbs = TRUE, summaryFunction = multiClassSummary, savePredictions = TRUE, verboseIter = TRUE)
  nocctrl <- trainControl(method="none",classProbs = TRUE)
  p="EarlyRec"
  fm <- as.formula( paste( p, ".", sep=" ~ ")) 
  rm(p)
}
Finalmodels<-qread('Finalmodels4.1.q')
```


# Model Training

To train the models, first the hyperparameter grids are defined, before training parameters are set.

<br><br>
```{r cars,echo=TRUE,eval=FALSE}

###Grids & TR control
{
ELGrid = expand.grid(alpha = seq(0,1,by=0.1),lambda = seq(0,0.1,by = 0.005))
RFGrid = expand.grid(
  .mtry=2:10,
  .splitrule=c("gini","extratrees"),
  .min.node.size=c(1,3,5)
)
XGBGrid <- expand.grid(
  nrounds = seq(from = 0, to = 1000, by = 200),
  eta = c(0.05, 0.1, 0.3),
  max_depth = c(2, 4, 6),
  gamma = c(0.5,0.7,1.0),
  colsample_bytree = c(0.4,0.8),
  min_child_weight = c(1,2,3),
  subsample = c(0.5,0.75))
}
  cctrl1 <- trainControl(method="repeatedcv", number=10,returnResamp = "all", repeats=5, 
                         classProbs = TRUE, summaryFunction = multiClassSummary, savePredictions = TRUE, verboseIter = TRUE,allowParallel = TRUE)
  cctrl3 <- trainControl(method="boot", number=1000,returnResamp = "all", 
                         classProbs = TRUE, summaryFunction = multiClassSummary, savePredictions = TRUE, verboseIter = TRUE)
  nocctrl <- trainControl(method="none",classProbs = TRUE)
  p="EarlyRec"
  fm <- as.formula( paste( p, ".", sep=" ~ ")) 
  rm(p)
}
```

<br><br>
Each model is then fitted and the hyperparameter combination that yields the highest mean AUC across 5 repeats of 10-fold cross validation is chosen as the optimum hyperparameter. This can then be validated using 1000 bootstrap resamples as below. Note the importance='permutation' argument that is given to the Random Forest model, which is required for variable importance calculations later.
<br><br>
```{r cars2,echo=TRUE,eval=FALSE}
####Elastic Net

registerDoParallel(cores=3)
system.time(ELRFull<- train(fm, data=dataC, method = "glmnet",preProcess=c('knnImpute'), trControl = cctrl1,
                            metric = "AUC",na.action=na.pass,tuneGrid = ELGrid))
system.time(ELRFinal<-train(fm, data=dataC, method = "glmnet",preProcess=c('knnImpute'), trControl = cctrl3,
                            metric = "AUC",na.action=na.pass,tuneGrid = expand.grid(alpha = ELRFull$bestTune$alpha,lambda = ELRFull$bestTune$lambda)))
registerDoSEQ()

saveRDS(ELRFull,file="ELRFull4.1.rds")
saveRDS(ELRFinal,file="ELRFinal4.1.rds")

####Random Forest
  registerDoParallel(cores=3)
  
  system.time(RFFull<-train(fm, data=dataC, method="ranger", num.trees=1000,preProcess=c('knnImpute'),
                            na.action=na.pass,replace=TRUE, trControl= cctrl1, tuneGrid=RFGrid, metric ="AUC"))

  system.time(RFFinal<-train(fm, data=dataC, method="ranger", num.trees=1000,preProcess=c('knnImpute'),
                             na.action=na.pass,replace=TRUE,importance='permutation',trControl= cctrl3,
                             tuneGrid=RFFull$bestTune, metric ="AUC"))
  registerDoSEQ()
  saveRDS(RFFull,'RFFull4.1rds')
  saveRDS(RFFinal,file="RFFinal4.1.rds")

#####XGBoost

  registerDoParallel(cores=3)
  XGBFull <-train(fm,data=dataC,preProcess=c('knnImpute'),method="xgbTree",trControl=cctrl1,tuneGrid=XGBGrid,
                  verbose=T,metric="AUC",na.action = na.pass)
  final_grid<-XGBFull$bestTune
  
  XGBFinal<-train(fm,data=dataC,preProcess=c('knnImpute'),method="xgbTree",trControl=cctrl3,tuneGrid=final_grid,
                  verbose=T,metric="AUC",na.action = na.pass)
  registerDoSEQ()
saveRDS(XGBFull,'XGBFull4.1.rds')
saveRDS(XGBFinal,'XGBFinal4.1.rds')


#####Ensemble
{
  ensboot<-createResample(dataC$EarlyRec,times=1)
  enstrain1<-dataC[ensboot[[1]],]
  enstrain2<-dataC[-ensboot[[1]],]
  registerDoParallel(cores=3)
  ELRFull2<- train(EarlyRec~., data=dataC, method = "glmnet",preProcess=c('knnImpute'), trControl = cctrl1,
                   metric = "AUC",na.action=na.pass,tuneGrid = ELGrid)
  ELRFinal2<- train(EarlyRec~., data=dataC, method = "glmnet",preProcess=c('knnImpute'), trControl = nocctrl,
                    metric = "AUC",na.action=na.pass,tuneGrid = ELRFull2$bestTune)
  qsave(ELRFinal2,'ELRFinal2ens.q')
  RFFull2<-train(EarlyRec~., data=dataC, method="ranger", num.trees=1000,preProcess=c('knnImpute') ,na.action=na.pass,replace=TRUE,
                            trControl= cctrl1, tuneGrid=RFGrid, metric ="AUC")
  RFFinal2<-train(EarlyRec~., data=dataC, method="ranger", num.trees=1000,preProcess=c('knnImpute') ,na.action=na.pass,replace=TRUE,
                 trControl= nocctrl, tuneGrid=RFFull2$bestTune, metric ="AUC")
  qsave(RFFinal2,'RFFinal2ens.q')
  system.time(XGBFull2 <-train(EarlyRec~.,data=dataC,preProcess=c('knnImpute'),method="xgbTree",trControl=cctrl1,tuneGrid=XGBGrid,
                   verbose=T,metric="AUC",na.action = na.pass))
  XGBFinal2 <-train(EarlyRec~.,data=dataC,preProcess=c('knnImpute'),method="xgbTree",trControl=nocctrl,tuneGrid=XGBFull2$bestTune,
                   verbose=T,metric="AUC",na.action = na.pass)
  qsave(XGBFinal2,'XGBFinal2ens.q')
  registerDoSEQ()
  rm(ELRFull2,RFFull2,XGBFull2)
  RFFinal2<-qread('RFFinal2ens.q')
  ELRFinal2<-qread('ELRFinal2ens.q')
  preds<-as.data.frame(1:nrow(na.omit(enstrain2)))
  preds$ELR<-predict(ELRFinal2,enstrain2,type='prob')[,'Yes']
  preds$RF<-predict(RFFinal2,enstrain2,type='prob')[,'Yes']
  preds$XGB<-predict(XGBFinal2,enstrain2,type='prob')[,'Yes']
  preds$Obs<-na.omit(enstrain2)$EarlyRec
  preds<-select(preds,-1)
  EnsLRM<-train(Obs~., data=preds, method = "glm",preProcess=c('knnImpute'), trControl = nocctrl,
        metric = "AUC",na.action=na.pass)
  EnsembleFinal<-list(ELRFinal2,RFFinal2,XGBFinal2,EnsLRM)

  Enspredfct=function(model,x){
    preds<-as.data.frame(predict(model[[4]][[1]],x,type='prob')[,'Yes'])
    colnames(preds)<-'ELR'
    preds$RF<-predict(model[[4]][[2]],x,type='prob')[,'Yes']
    preds$XGB<-predict(model[[4]][[3]],x,type='prob')[,'Yes']
    preds$LRM<-predict(model[[4]][[4]],preds,type='prob')[,'Yes']
    preds$LRM
  }
}
```


<br><br>
The ensemble model needs to be trained slightly differently, because of the need for the linear blending of probabilities. The dataset is first split into in-bootstrap and out-bootstrap samples, then the base models are trained as above on the in-bootstrap samples. These models are then used to predict on the out-boostrap samples, which are then combined together as a linear blend using logistic regression. The predict function defined below ('Enspredfct') is then used to generate predictions from the 'Ensemblefinal' object.
<br><br>
```{r cars3,echo=TRUE,eval=FALSE}
#####Ensemble
{
  ensboot<-createResample(dataC$EarlyRec,times=1)
  enstrain1<-dataC[ensboot[[1]],]
  enstrain2<-dataC[-ensboot[[1]],]
  registerDoParallel(cores=3)
  ELRFull2<- train(EarlyRec~., data=dataC, method = "glmnet",preProcess=c('knnImpute'), trControl = cctrl1,
                   metric = "AUC",na.action=na.pass,tuneGrid = ELGrid)
  ELRFinal2<- train(EarlyRec~., data=dataC, method = "glmnet",preProcess=c('knnImpute'), trControl = nocctrl,
                    metric = "AUC",na.action=na.pass,tuneGrid = ELRFull2$bestTune)
  qsave(ELRFinal2,'ELRFinal2ens.q')
  RFFull2<-train(EarlyRec~., data=dataC, method="ranger", num.trees=1000,preProcess=c('knnImpute') ,na.action=na.pass,replace=TRUE,
                            trControl= cctrl1, tuneGrid=RFGrid, metric ="AUC")
  RFFinal2<-train(EarlyRec~., data=dataC, method="ranger", num.trees=1000,preProcess=c('knnImpute') ,na.action=na.pass,replace=TRUE,
                 trControl= nocctrl, tuneGrid=RFFull2$bestTune, metric ="AUC")
  qsave(RFFinal2,'RFFinal2ens.q')
  system.time(XGBFull2 <-train(EarlyRec~.,data=dataC,preProcess=c('knnImpute'),method="xgbTree",trControl=cctrl1,tuneGrid=XGBGrid,
                   verbose=T,metric="AUC",na.action = na.pass))
  XGBFinal2 <-train(EarlyRec~.,data=dataC,preProcess=c('knnImpute'),method="xgbTree",trControl=nocctrl,tuneGrid=XGBFull2$bestTune,
                   verbose=T,metric="AUC",na.action = na.pass)
  qsave(XGBFinal2,'XGBFinal2ens.q')
  registerDoSEQ()
  rm(ELRFull2,RFFull2,XGBFull2)
  RFFinal2<-qread('RFFinal2ens.q')
  ELRFinal2<-qread('ELRFinal2ens.q')
  preds<-as.data.frame(1:nrow(na.omit(enstrain2)))
  preds$ELR<-predict(ELRFinal2,enstrain2,type='prob')[,'Yes']
  preds$RF<-predict(RFFinal2,enstrain2,type='prob')[,'Yes']
  preds$XGB<-predict(XGBFinal2,enstrain2,type='prob')[,'Yes']
  preds$Obs<-na.omit(enstrain2)$EarlyRec
  preds<-select(preds,-1)
  EnsLRM<-train(Obs~., data=preds, method = "glm",preProcess=c('knnImpute'), trControl = nocctrl,
        metric = "AUC",na.action=na.pass)
  EnsembleFinal<-list(ELRFinal2,RFFinal2,XGBFinal2,EnsLRM)

  Enspredfct=function(model,x){
    preds<-as.data.frame(predict(model[[4]][[1]],x,type='prob')[,'Yes'])
    colnames(preds)<-'ELR'
    preds$RF<-predict(model[[4]][[2]],x,type='prob')[,'Yes']
    preds$XGB<-predict(model[[4]][[3]],x,type='prob')[,'Yes']
    preds$LRM<-predict(model[[4]][[4]],preds,type='prob')[,'Yes']
    preds$LRM
  }
}
```


# Internal validation by the double bootstrap

Similar to the above, the data is first split into in/out of bootstrap samples (the outer bootstrap,500 times). Each model is then trained on the outerbootstrap training samples, and hyperparameters optimised by an inner bootstrap (1000 times). The same inner and outer bootstrap cases are used for each modelling strategy to ensure they can be combined and compared easily. The inner bootstrap resamples are then extracted and isotonic and logistic regression (for the ensemble) models are generated using the inner bootstrap testing resamples.

The base models and the isotonic and logistic regression models can then be applied to the outer bootstrap testing data to obtain performance metrics. For each outer bootstrap resample, the 1000 inner bootstrap resample outputs are averaged, and then averaged across all the outer boostrap resamples for the final metrics and plots.
<br><br>
```{r cars4,echo=TRUE,eval=FALSE}

###Outerbootstrap
  part<-createResample(dataC$EarlyRec,times=500)
  outerbootdata<-1
  ####Innerbootstrap
 system.time(for (p in 1:500){
  trainpart<-dataC[part[[p]],]
  testpart<-dataC[-part[[p]],]
  boot<-createResample(trainpart$EarlyRec,times=1000)
  cctrl3x <- trainControl(index=boot,method="boot", number=1000,returnResamp = "all", 
                          classProbs = TRUE, summaryFunction = multiClassSummary, savePredictions = TRUE, verboseIter = TRUE)

####Train models
  registerDoParallel(cores=3)
ELR<-train(EarlyRec~., data=trainpart,preProcess=c('knnImpute'),method = "glmnet", trControl = cctrl3x,
           metric = "AUC",na.action=na.pass,tuneGrid = Finalmodels[[1]]$bestTune)
RF<-train(EarlyRec~., data=trainpart,preProcess=c('knnImpute'), method="ranger", num.trees=1000, 
          na.action=na.pass,replace=TRUE,importance='none',trControl= cctrl3x, 
          tuneGrid=Finalmodels[[2]]$bestTune, metric ="AUC")
XGB<-train(EarlyRec~.,data=trainpart,preProcess=c('knnImpute'),method="xgbTree",
           trControl=cctrl3x,tuneGrid=Finalmodels[[3]]$bestTune,verbose=T,metric="AUC",na.action = na.pass)
    registerDoSEQ()

    cabs<-1
    system.time(for (z in 1:length(levels(as.factor(ELR$pred$Resample)))){
      ELR$pred$Resample<-as.factor(ELR$pred$Resample)
      RF$pred$Resample<-as.factor(RF$pred$Resample)
      XGB$pred$Resample<-as.factor(XGB$pred$Resample)
    x1<-ELR$pred[ELR$pred$Resample==levels(ELR$pred$Resample)[z],]
    x2<-RF$pred[RF$pred$Resample==levels(RF$pred$Resample)[z],]
    x3<-XGB$pred[XGB$pred$Resample==levels(XGB$pred$Resample)[z],]
    preds<-as.data.frame(cbind(x1$Yes,x2$Yes,x3$Yes,as.numeric(x1$obs)-1))
    colnames(preds)<-c('ELR','RF','XGB','obs')
    elrcal<-CORElearn::calibrate(as.factor(preds$obs),preds$ELR,class1=2,method='isoReg',assumeProbabilities = TRUE)
    rfcal<-CORElearn::calibrate(as.factor(preds$obs),preds$RF,class1=2,method='isoReg',assumeProbabilities = TRUE)
    xgbcal<-CORElearn::calibrate(as.factor(preds$obs),preds$XGB,class1=2,method='isoReg',assumeProbabilities = TRUE)
    preds$ELRc<-applyCalibration(preds$ELR,elrcal)
    preds$RFc<-applyCalibration(preds$RF,rfcal)
    preds$XGBc<-applyCalibration(preds$XGB,xgbcal)
    lrm<-lrm(obs~ELR+RF+XGB,data=preds)
    lrmc<-lrm(obs~ELRc+RFc+XGBc,data=preds)
    preds$LRM<-predict(lrm,preds,type='fitted')
    lrmc2<-lrm(obs~LRM,data=preds)
    preds$LRMc<-predict(lrmc,preds,type='fitted')
    cabs[z]<-list(list(elrcal,rfcal,xgbcal,lrm,lrmc,lrmc2))
})####30s
    
  predictions<-1  
  system.time(  for (z in 1:length(levels(as.factor(ELR$pred$Resample)))){
      elrcal<-cabs[[z]][[1]]
      rfcal<-cabs[[z]][[2]]
      xgbcal<-cabs[[z]][[3]]
      lrm<-cabs[[z]][[4]]
      lrmc<-cabs[[z]][[5]]
      lrmc2<-cabs[[z]][[6]]
      
preds2<-as.data.frame(predict(ELR,testpart,type='prob')[,'Yes'])
colnames(preds2)<-'ELR'
preds2$RF<-predict(RF,testpart,type='prob')[,'Yes']
preds2$XGB<-predict(XGB,testpart,type='prob')[,'Yes']
preds2$LRM<-predict(lrm,preds2,type='fitted')
preds2$ELRc<-applyCalibration(preds2$ELR,elrcal)
preds2$RFc<-applyCalibration(preds2$RF,rfcal)
preds2$XGBc<-applyCalibration(preds2$XGB,xgbcal)
preds2$LRMc<-predict(lrmc,preds2,type='fitted')
preds2$LRMc2<-predict(lrmc2,preds2,type='fitted')
preds2$obs<-as.numeric(na.omit(testpart)$EarlyRec)-1
predictions[z]<-list(preds2)
})#####64s
 
 system.time({ 
   avefoldLL<-1
  avefoldBR<-1
  avefoldCID<-1
  calplot1<-1
  ROCplot1<-1
 cabdata2<-1
 rocdata2<-1
 for (i in 1:4){
   cabdata<-1
   brs<-1
   LLs<-1
   cids<-1
   lgt<-1
   roc<-1
   for (z in 1:1){
     df<-as.data.frame(cbind(predictions[[z]][,i],predictions[[z]][,10]))
     colnames(df)<-c('Preds','Obs')
     brs[z]<-BrierScore(df$Obs,df$Preds)
     LLs[z]<-logLoss(df$Obs,df$Preds)
     cids[z]<-auc(df$Obs,df$Preds)
     lgt[z]<-nrow(df)
     
     xx<-ggplot_build(ggplot(data = df) + xlab("Predicted Probability") + ylab('Observed Average')+
                        geom_smooth(data = df, aes(Preds, Obs),
                                    color = "#F8766D",se=TRUE,level=0.95,span=2/3)+xlim(c(0,1))+ylim(c(0,1))+
                        geom_abline(intercept = 0, slope = 1,linetype=2))
     for (m in 1:(length(xx$data[[1]]$y)-1)){
       xx$data[[1]]$y[m+1]<-ifelse(xx$data[[1]]$y[m]>0.9&is.na(xx$data[[1]]$y[m+1]),1.0,
                                   ifelse(xx$data[[1]]$y[m]<0.1&is.na(xx$data[[1]]$y[m+1]),0,xx$data[[1]]$y[m+1]))
       
     }
     cabdata[z]<-list(select(xx$data[[1]],x,y))
     rocobj<-roc(df$Obs,df$Preds)
     roc[z]<-list(ggplot_build(ggroc(rocobj,alpha=0.5,colour='red',linetype=6,size=1))$data)
   }
   avefoldLL[i]<-list(c(mean(LLs),std.error(LLs)))
   avefoldBR[i]<-list(c(mean(brs),std.error(brs)))
   avefoldCID[i]<-list(c(mean(cids),std.error(cids)))
   
   lgt2<-1
   for (n in 1:length(roc)){
     lgt2[n]<-length(roc[[n]][[1]]$x)
   }
   xR<-as.data.frame(roc[[1]][[1]]$x[(1:min(lgt2))])
   yR<-as.data.frame(roc[[1]][[1]]$y[(1:min(lgt2))])
   for (n in 1:length(roc)){
     seqs<-ceiling(seq(from=1,to=length(roc[[n]][[1]]$x),length.out=min(lgt2)))
     xR[,n]<-roc[[n]][[1]]$x[seqs]
     yR[,n]<-roc[[n]][[1]]$y[seqs]
   }
   rocdat<-as.data.frame(cbind(-rowMeans(xR),
                               rowMeans(yR),
                               rowMeans(yR)+1.96*std.error(t(yR)),
                               rowMeans(yR)-1.96*std.error(t(yR))))
   
   colnames(rocdat)<-c('x','y','ymax','ymin')
   ROCplot1[i]<-list(ggplot(data = rocdat) + xlab("Specificity") + ylab('Sensitivity')+
                         geom_line(data = rocdat, aes(y=y, x=x),
                                   color = "#F8766D")+xlim(c(1,0))+ylim(c(0,1))+
                         geom_abline(intercept = 1, slope = 1,linetype=2)+
                         geom_ribbon(data=rocdat,aes(x=x,ymin=ymin,ymax=ymax), fill="grey", alpha=0.25))
   
   yy<-ggplot_build(ggplot(data = rocdat) + xlab("Specificity") + ylab('Sensitivity')+
                      geom_line(data = rocdat, aes(y=y, x=x),
                                color = "#F8766D")+xlim(c(1,0))+ylim(c(0,1))+
                      geom_abline(intercept = 1, slope = 1,linetype=2)+
                      geom_ribbon(data=rocdat,aes(x=x,ymin=ymin,ymax=ymax), fill="grey", alpha=0.25))
   
   rocdata2[i]<-list(as.data.frame(c(select(yy$data[[1]],x,y),select(yy$data[[3]],ymax,ymin))))
   xC<-as.data.frame(cabdata[[1]]$x)
   yC<-as.data.frame(cabdata[[1]]$y)
   for (n in 1:length(cabdata)){
     xC[,n]<-cabdata[[n]]$x
     yC[,n]<-cabdata[[n]]$y
   }
   yC[is.na(yC)]<-0
   cal_plot_data2<-cabdata[[1]]
   cal_plot_data2$x<-rowMeans(xC)
   cal_plot_data2$y<-rowMeans(yC)
   cal_plot_data2$ymin<-rowMeans(yC)-1.96*std.error(t(yC))
   cal_plot_data2$ymax<-rowMeans(yC)+1.96*std.error(t(yC))
   calplot1[i]<-list(ggplot(data = cal_plot_data2) + xlab("Predicted Probability") + ylab('Observed Average')+
                         geom_smooth(data = cal_plot_data2, aes(x, y),
                                     color = "#F8766D",se=FALSE,level=0.95,span=2/3)+xlim(c(0,1))+ylim(c(0,1))+
                         geom_abline(intercept = 0, slope = 1,linetype=2)+
                         geom_ribbon(data=cal_plot_data2,aes(x=x,ymin=ymin,ymax=ymax), fill="grey", alpha=0.25))
   yy<-ggplot_build(ggplot(data = cal_plot_data2) + xlab("Predicted Probability") + ylab('Observed Average')+
                      geom_smooth(data = cal_plot_data2, aes(x, y),
                                  color = "#F8766D",se=FALSE,level=0.95,span=2/3)+xlim(c(0,1))+ylim(c(0,1))+
                      geom_abline(intercept = 0, slope = 1,linetype=2)+
                      geom_ribbon(data=cal_plot_data2,aes(x=x,ymin=ymin,ymax=ymax), fill="grey", alpha=0.25))
   cabdata2[i]<-list(as.data.frame(c(select(yy$data[[1]],x,y),select(yy$data[[3]],ymax,ymin))))
 }
  for (i in 5:9){
    cabdata<-1
    brs<-1
    LLs<-1
    cids<-1
    lgt<-1
    roc<-1
    for (z in 1:length(predictions)){
      df<-as.data.frame(cbind(predictions[[z]][,i],predictions[[z]][,10]))
      colnames(df)<-c('Preds','Obs')
      brs[z]<-BrierScore(df$Obs,df$Preds)
      LLs[z]<-logLoss(df$Obs,df$Preds)
      cids[z]<-auc(df$Obs,df$Preds)
      lgt[z]<-nrow(df)
      
      xx<-ggplot_build(ggplot(data = df) + xlab("Predicted Probability") + ylab('Observed Average')+
                         geom_smooth(data = df, aes(Preds, Obs),
                                     color = "#F8766D",se=TRUE,level=0.95,span=2/3)+xlim(c(0,1))+ylim(c(0,1))+
                         geom_abline(intercept = 0, slope = 1,linetype=2))
      if(nrow(xx$data[[1]])>0){
      for (m in 1:(length(xx$data[[1]]$y)-1)){
        xx$data[[1]]$y[m+1]<-ifelse(xx$data[[1]]$y[m]>0.9&is.na(xx$data[[1]]$y[m+1]),1.0,
                                    ifelse(xx$data[[1]]$y[m]<0.1&is.na(xx$data[[1]]$y[m+1]),0,xx$data[[1]]$y[m+1]))
        
      }
      }else
      {
        xx<-ggplot_build(ggplot(data = df) + xlab("Predicted Probability") + ylab('Observed Average')+
                           geom_smooth(data = df, aes(Preds, Obs),
                                       color = "#F8766D",se=TRUE,level=0.95,span=1)+xlim(c(0,1))+ylim(c(0,1))+
                           geom_abline(intercept = 0, slope = 1,linetype=2))
      for (m in 1:(length(xx$data[[1]]$y)-1)){
        xx$data[[1]]$y[m+1]<-ifelse(xx$data[[1]]$y[m]>0.9&is.na(xx$data[[1]]$y[m+1]),1.0,
                                    ifelse(xx$data[[1]]$y[m]<0.1&is.na(xx$data[[1]]$y[m+1]),0,xx$data[[1]]$y[m+1]))
        
      }
      }
      cabdata[z]<-list(select(xx$data[[1]],x,y))
      rocobj<-roc(df$Obs,df$Preds)
      roc[z]<-list(ggplot_build(ggroc(rocobj,alpha=0.5,colour='red',linetype=6,size=1))$data)
    }
    avefoldLL[i]<-list(c(mean(LLs),std.error(LLs)))
    avefoldBR[i]<-list(c(mean(brs),std.error(brs)))
    avefoldCID[i]<-list(c(mean(cids),std.error(cids)))
    lgt2<-1
    for (n in 1:length(roc)){
      lgt2[n]<-length(roc[[n]][[1]]$x)
    }
    xR<-as.data.frame(roc[[1]][[1]]$x[(1:min(lgt2))])
    yR<-as.data.frame(roc[[1]][[1]]$y[(1:min(lgt2))])
    for (n in 1:length(roc)){
      seqs<-ceiling(seq(from=1,to=length(roc[[n]][[1]]$x),length.out=min(lgt2)))
      xR[,n]<-roc[[n]][[1]]$x[seqs]
      yR[,n]<-roc[[n]][[1]]$y[seqs]
    }
    rocdat<-as.data.frame(cbind(-rowMeans(xR),
                                rowMeans(yR),
                                rowMeans(yR)+1.96*std.error(t(yR)),
                                rowMeans(yR)-1.96*std.error(t(yR))))
    
    colnames(rocdat)<-c('x','y','ymax','ymin')
    ROCplot1[i]<-list(ggplot(data = rocdat) + xlab("Specificity") + ylab('Sensitivity')+
                          geom_line(data = rocdat, aes(y=y, x=x),
                                    color = "#F8766D")+xlim(c(1,0))+ylim(c(0,1))+
                          geom_abline(intercept = 1, slope = 1,linetype=2)+
                          geom_ribbon(data=rocdat,aes(x=x,ymin=ymin,ymax=ymax), fill="grey", alpha=0.25))
    
    yy<-ggplot_build(ggplot(data = rocdat) + xlab("Specificity") + ylab('Sensitivity')+
                        geom_line(data = rocdat, aes(y=y, x=x),
                                  color = "#F8766D")+xlim(c(1,0))+ylim(c(0,1))+
                        geom_abline(intercept = 1, slope = 1,linetype=2)+
                        geom_ribbon(data=rocdat,aes(x=x,ymin=ymin,ymax=ymax), fill="grey", alpha=0.25))
    rocdata2[i]<-list(as.data.frame(c(select(yy$data[[1]],x,y),select(yy$data[[3]],ymax,ymin))))
    avefoldLL  
    avefoldBR
    xC<-as.data.frame(cabdata[[1]]$x)
      yC<-as.data.frame(cabdata[[1]]$y)
      for (n in 1:length(cabdata)){
        xC[,n]<-cabdata[[n]]$x
        yC[,n]<-cabdata[[n]]$y
      }
      yC[is.na(yC)]<-0
      cal_plot_data2<-cabdata[[1]]
      cal_plot_data2$x<-rowMeans(xC)
      cal_plot_data2$y<-rowMeans(yC)
      cal_plot_data2$ymin<-rowMeans(yC)-1.96*std.error(t(yC))
      cal_plot_data2$ymax<-rowMeans(yC)+1.96*std.error(t(yC))
      calplot1[i]<-list(ggplot(data = cal_plot_data2) + xlab("Predicted Probability") + ylab('Observed Average')+
        geom_smooth(data = cal_plot_data2, aes(x, y),
                    color = "#F8766D",se=FALSE,level=0.95,span=2/3)+xlim(c(0,1))+ylim(c(0,1))+
        geom_abline(intercept = 0, slope = 1,linetype=2)+
        geom_ribbon(data=cal_plot_data2,aes(x=x,ymin=ymin,ymax=ymax), fill="grey", alpha=0.25))
      yy<-ggplot_build(ggplot(data = cal_plot_data2) + xlab("Predicted Probability") + ylab('Observed Average')+
        geom_smooth(data = cal_plot_data2, aes(x, y),
                    color = "#F8766D",se=FALSE,level=0.95,span=2/3)+xlim(c(0,1))+ylim(c(0,1))+
        geom_abline(intercept = 0, slope = 1,linetype=2)+
        geom_ribbon(data=cal_plot_data2,aes(x=x,ymin=ymin,ymax=ymax), fill="grey", alpha=0.25))
      cabdata2[i]<-list(as.data.frame(c(select(yy$data[[1]],x,y),select(yy$data[[3]],ymax,ymin))))
  }

 names(avefoldLL)<-colnames(predictions[[1]])[1:9]
 names(avefoldBR)<-colnames(predictions[[1]])[1:9]
 names(avefoldCID)<-colnames(predictions[[1]])[1:9]
 names(calplot1)<-colnames(predictions[[1]])[1:9]
 names(ROCplot1)<-colnames(predictions[[1]])[1:9]
 names(cabdata2)<-colnames(predictions[[1]])[1:9]
 names(rocdata2)<-colnames(predictions[[1]])[1:9]
  })
outerbootdata[p]<-list(list(avefoldLL,avefoldBR,avefoldCID,calplot1,ROCplot1,cabdata2,rocdata2))
 names(outerbootdata[[p]])<-c('LogLoss','Brier','C-index','Calibration Plots','ROC Plots','Calibration Data','ROC Data')
 qsave(outerbootdata,'outerbootdata.q')

}  

```
<br><br>

# Variable Importance Partial Dependence

The partial dependence function is derived below with two example variables (age and gender). The prediction function for the enemble model is defined as earlier (i.e. 'Enspredfct' ). Raw values can be extracted from the 'ft' object if required.


```{r cars5,echo=TRUE,message=FALSE,warning=FALSE}
dataNA<-na.omit(dataC)

expELR<-explain(Finalmodels[[1]],data=dataNA,y=as.numeric(dataNA$EarlyRec)-1,label='ELR')
expRF<-explain(Finalmodels[[2]],data=dataNA,y=as.numeric(dataNA$EarlyRec)-1,label='RF')
expXGB<-explain(Finalmodels[[3]],data=dataNA,y=as.numeric(dataNA$EarlyRec)-1,label='XGBoost')
expENS<-explain(Finalmodels,data=dataNA,y=as.numeric(dataNA$EarlyRec)-1,label='Ensemble',predict_function=Finalmodels[[5]])

x<-(ingredients::partial_dependency(expXGB,'Age'))
x1<-(ingredients::partial_dependency(expRF,'Age'))
x2<-(ingredients::partial_dependency(expELR,'Age'))
x3<-(ingredients::partial_dependency(expENS,'Age'))
fintab<-rbind(x,x1,x2,x3)
ft<-rbind(x,x1,x2,x3)
colnames(fintab)<-c('Variable.Name','Model','x','y','ids')
agepdp<-ggplotly(ggplot(fintab,aes(x,y))+
                   geom_line(data=fintab,mapping=aes(colour=Model))+theme_bw()+
                   ylim(0,0.6)+
                   ylab('Predicted Probability')+
                   xlab('Age (years)')+ggtitle('Age at diagnosis'))
agepdp

x<-(ingredients::partial_dependency(expXGB,'Gender'))
x1<-(ingredients::partial_dependency(expRF,'Gender'))
x2<-(ingredients::partial_dependency(expELR,'Gender'))
x3<-(ingredients::partial_dependency(expENS,'Gender'))
fintab<-rbind(x,x1,x2,x3)
ft<-rbind(ft,fintab)
colnames(fintab)<-c('Variable.Name','Model','x','y','ids')
colnames(fintab)[3]<-'Gender'
Genderpdp<-ggplot(fintab,aes(x=Gender,y=y,fill=Gender))+
  geom_bar(position='dodge',stat='identity')+theme_bw()+
  ylab('Marginal Predicted Probability')+
  xlab('')+
  facet_wrap(~Model)+ggtitle('Gender')
Genderpdp

```
